\chapter{Implementation}
\label{ch:implementation}

\section{System Architecture}
\label{sec:system-architecture}

\subsection{Modular Design}
\label{sec:modular-design}

Our implementation follows a modular design with separate components for data
loading, feature extraction, model training, and evaluation.

\subsection{Configuration Management}
\label{sec:config-management}

All parameters are centralised in \texttt{config.py} for easy experimentation:
\begin{itemize}
  \item Dataset paths and model names
  \item Hardware settings (CPU/GPU)
  \item Evaluation thresholds (NED, IoU)
  \item Training hyperparameters
  \item Clustering parameters ($k$, random seed)
\end{itemize}

\subsection{Modular Pipeline Architecture}
\label{sec:pipeline-architecture}

The implementation follows a staged pipeline design:
\begin{enumerate}
  \item Stage~0: Evaluation protocol definition
  \item Stage~1: Dataset ingestion and preprocessing
  \item Stage~2: Layout clustering and analysis
  \item Stage~3: Baseline implementation and evaluation
  \item Stage~4: Main model (LayoutLMv3\,+\,Linker)
\end{enumerate}
Each stage is implemented as an independent function in \texttt{main.py},
allowing for incremental development, easy debugging, flexible execution, and
reproducible experiments with consistent interfaces.

% ---------------------------------------------------------------------------
\section{Data Preprocessing}
\label{sec:data-preprocessing}

The data preprocessing pipeline was extended substantially during the thesis
because the raw KVP10k rows were not directly usable for the Mistral baseline.
In particular, the learned baseline required page-level document text, bounding
boxes, and aligned key-value pair supervision in a format consistent with the
intended IBM-style prompt and target representation.

\subsection{Image Processing}
\label{sec:image-processing}

For the Mistral baseline, preprocessing begins from the original PDF source
referenced by each dataset item. The relevant page is accessed using the
dataset page number and converted from the dataset's one-indexed page numbering
to PyMuPDF's zero-indexed page access. Page size is preserved and later mapped
to a quantised $100 \times 100$ coordinate system for prompt and target
generation.

For Stage~4, pages are additionally rendered at 150\,DPI and resized to
$224 \times 224$ before being passed to the LayoutLMv3 image encoder.
The resize preserves the document's aspect ratio up to the nearest 14-pixel
boundary (matching the $14 \times 14$ patch grid) before centre-cropping.

\subsection{Text Processing}
\label{sec:text-processing}

Instead of relying on an external OCR engine, the implementation uses PyMuPDF
to extract word-level text and bounding boxes directly from the PDF text layer.
This choice was made for practical deployment reasons on the cluster
environment, where adding a system OCR dependency would have complicated
execution. The trade-off is that scanned pages without a text layer cannot
contribute usable words and are therefore discarded during preparation.

The extracted words are grouped into line-like units and converted into an
LMDX-style textual document representation. This representation is later
embedded into the prompt given to the Mistral model. Ground-truth targets are
serialised as Python list-of-lists strings, following the benchmark style used
by the corrected Stage~3 pipeline.

\paragraph{V4: Word-level splitting.}
Diagnostic analysis (Section~\ref{sec:diagnostic-experiments}) revealed that
the LMDX text format stores entire OCR text blocks --- often multi-word
phrases such as \texttt{"1~lb.\ shelled \$15.00"} --- as single ``words''
with a single bounding box. This is problematic for LayoutLMv3, which expects
individual words, and for ground-truth label generation, because a key-value
pair whose key and value text reside within the same OCR block cannot produce
distinct entity labels or link labels at the block level.

The V4 data pipeline splits each OCR block into whitespace-separated words and
estimates horizontal sub-bounding boxes proportional to character count. For
example, the block \texttt{"1~lb.\ shelled \$15.00"} with bounding box
$[x_1, y_1, x_2, y_2]$ is split into four words (\texttt{"1"},
\texttt{"lb."}, \texttt{"shelled"}, \texttt{"\$15.00"}), each assigned a
horizontal slice of the original bbox. This increased the average word count
per document from $\sim$35 (block-level) to $\sim$200 (word-level) and
enabled the entity labels and link labels to reach word-level granularity.

\subsection{Layout Encoding}
\label{sec:layout-encoding}

Annotation geometry is fused with extracted words through an OCR-to-annotation
matching step. Bounding boxes are converted into normalised integer coordinates
on a $100 \times 100$ grid. The resulting aligned entities are then transformed
into three categories: regular key-value pairs, unkeyed values, and unvalued
keys. This preprocessing stage also handles repeated annotator copies in
KVP10k by grouping rows by page identifier and selecting the richest annotation
set for each unique page.

For Stage~4, bounding box coordinates are further re-normalised to the
$1000 \times 1000$ LayoutLMv3 convention (integers in $[0, 1000]$) after the
above deduplication step.

% ---------------------------------------------------------------------------
\section{Model Implementation}
\label{sec:model-implementation}

\subsection{LayoutLMv3 Integration}
\label{sec:layoutlmv3-integration}

The Stage~4 model class \texttt{LayoutLMv3KVPModel} wraps the
\texttt{microsoft/layoutlmv3-base} checkpoint from the HuggingFace Hub.
The base encoder is loaded in full fp32 precision. No quantisation is applied:
the 125M-parameter LayoutLMv3-base model fits within the 40\,GB memory
envelope of the A100 nodes used for Stage~4, unlike the 7B-parameter Mistral
model that required QLoRA in Stage~3.

All encoder parameters are unfrozen and participate in fine-tuning. A linear
learning-rate warmup of 500 steps is used to avoid large gradient updates on
the pre-trained weights during the first portion of training (see
Section~\ref{sec:optimisation}).

Tokenisation follows the LayoutLMv3 tokenizer convention: each OCR word and
its bounding box are passed together, and sub-word tokens inherit the bounding
box of their parent word. The maximum sequence length is capped at 512 tokens;
longer pages are truncated from the right after the \texttt{[SEP]} position.

\subsection{Entity Classifier Head}
\label{sec:entity-classifier-impl}

The entity classification head is a two-layer MLP applied independently to
each text token output of the encoder (image patch tokens are excluded from
classification). The network maps the 768-dimensional token representation to
a 3-class logit vector (\texttt{O}, \texttt{KEY}, \texttt{VALUE}) as described
in Section~\ref{sec:entity-classifier}.

During training, the cross-entropy loss is computed over all text tokens.
Tokens corresponding to padding positions and to image patches are masked
from the loss. To address the class imbalance (the \texttt{O} class typically
represents $\sim$70\% of tokens), class weights are set to the inverse
frequency observed across a 1{,}000-sample calibration pass of the training
set at the start of each Stage~4a training run.

\subsection{Biaffine Linker Module}
\label{sec:linker-impl}

The linker module receives key and value representations extracted from the
entity classifier output. In V1, these are individual token hidden states;
in V2, they are span-level mean-pooled representations
(Section~\ref{sec:span-representation}).

Representations are augmented with 8 spatial features
(Section~\ref{sec:spatial-encoding}) before scoring. The semantic scorer uses
scaled dot-product attention (Section~\ref{sec:kv-attention}) --- an earlier
bilinear implementation was abandoned due to gradient explosion.

\paragraph{Teacher forcing.}
During training, the linker receives ground-truth entity labels for span
grouping rather than the entity classifier's \texttt{argmax} predictions.
This ensures the linker learns from correctly segmented spans. At inference
time, the linker receives predicted entity labels. This teacher-forcing
strategy decouples entity classification errors from linker training and is
standard practice in structured prediction~\cite{spade}.

\paragraph{NaN guard.}
Achieving stable training required a valid-label mask before the BCE call:
ground-truth link labels outside $[0, 1]$ (from uninitialised positions in
the adjacency matrix) are excluded.  This is analogous to the
\texttt{ignore\_index=-100} mechanism used for entity labels in
\texttt{CrossEntropyLoss}. The \texttt{pos\_weight} for the BCE loss is set
to $\min(n_{\text{neg}}/n_{\text{pos}},\, 50)$ in V1 and
$n_{\text{neg}}/n_{\text{pos}}$ (uncapped) in V2.

\paragraph{Memory efficiency.}
To avoid materialising large expanded tensors, scoring is performed in chunks
of 8 key rows. A hard cap of 128 key and 128 value candidates per page
(selected by highest entity-classifier confidence) bounds the worst-case
linker matrix to $128 \times 128$ in V1.

% ---------------------------------------------------------------------------
\section{Baseline Evaluation Pipeline}
\label{sec:baseline-eval-pipeline}

\subsection{Automated Evaluation Workflow}
\label{sec:eval-workflow}

Stage~3 implements a fully automated 5-step evaluation pipeline:
\begin{enumerate}
  \item \textbf{Ground Truth Conversion:} Convert HuggingFace annotations to
  IBM benchmark format (extract KV pairs from linking attributes, handle ID
  prefix matching for truncated identifiers, generate individual JSON files
  per document).
  \item \textbf{Prediction Generation:} Run baseline on test set (apply spatial
  pairing algorithm, save predictions in IBM-compatible format).
  \item \textbf{Benchmark Evaluation:} Automatically execute IBM's official
  benchmark (compute precision, recall, F1; evaluate text-only, location-only,
  and combined metrics).
  \item \textbf{Results Analysis:} Generate detailed statistics (type
  distribution, per-document statistics, sample examples).
  \item \textbf{Visualisation:} Create publication-ready plots at 300\,DPI.
\end{enumerate}

\subsection{Scalability}
\label{sec:scalability}

The pipeline is designed to scale from small validation sets to full datasets:
\begin{itemize}
  \item Quick validation: 20 samples ($\sim$25 seconds)
  \item Medium validation: 1{,}000 samples ($\sim$2--3 minutes)
  \item Full test set: 5{,}255 samples (scales linearly)
\end{itemize}

% ---------------------------------------------------------------------------
\section{Training Pipeline}
\label{sec:training-pipeline}

The training pipeline for the Mistral baseline (Stage~3) operates on the
prepared JSON files generated during preprocessing. Each sample contains the
prompt text, the expected target response, and the aligned ground-truth
key-value pairs. During tokenisation, prompt tokens are masked in the loss so
that optimisation is applied only to the response segment.

\subsection{Data Loader}
\label{sec:data-loader}

The data loader reads page-level JSON files from the prepared train and test
directories. Samples without prompt text or without non-empty supervision are
skipped to avoid training on degenerate examples. For Stage~4, a
\texttt{KVP10kEntityDataset} class wraps the same JSON files and additionally
loads the rendered page image from a pre-computed image cache on the cluster
filesystem. Image loading is performed lazily to minimise CPU-to-GPU transfer
latency.

\subsection{Training Loop}
\label{sec:stage4-training-loop}

\paragraph{Stage~3 (Mistral QLoRA).}
We used QLoRA with 4-bit NF4 quantisation due to memory constraints on the
available 40\,GB A100 GPU (see Section~\ref{sec:hardware} for a detailed
discussion of the implications).

\paragraph{Stage~4a -- Entity Classifier Pre-training.}
Stage~4a trains the LayoutLMv3 encoder together with the entity classification
head only, with the linker module disabled ($\lambda = 0$).
This pre-trains the representation before the more complex
linking objective is introduced. Training uses fp32 precision
on A100~80\,GB cluster nodes; no quantisation is required.
Gradient accumulation over 8 micro-steps gives an
effective batch size of 32. A SLURM job template with a wall-clock limit of
24~hours is used for this phase.

\paragraph{Stage~4b -- Joint Fine-tuning.}
Stage~4b resumes from the Stage~4a checkpoint and activates the biaffine
linker ($\lambda > 0$). The encoder learning rate is reduced by a factor of
10 relative to Stage~4a to avoid catastrophic forgetting of the entity
classification capability. Two variants are trained: V1 (token-level linking)
and V2 (span-level linking). A SLURM job with a 12-hour wall-clock limit
is used for each variant. Early stopping (patience~$= 3$) on validation
entity F1 typically terminates training within 10--15 epochs.

\paragraph{V4 link label generation.}
Diagnostic analysis of V2 training revealed that the ground-truth link label
matrix contained a cross-product explosion: for each key-value pair, \emph{all}
key-word indices were linked to \emph{all} value-word indices, rather than a
single representative pair. A document with 8 ground-truth KVPs produced
$\sim$1{,}049 positive link entries instead of~8, causing the span-level
collapse (Section~\ref{sec:v2-span-linking}) to mark nearly every span pair as
positive --- equivalent to no supervision at all.

The V4 fix assigns exactly one link per KVP: the first key word is linked to
the first value word, producing $N$ positive link entries for $N$ ground-truth
KVPs.  Together with the word-level splitting
(Section~\ref{sec:text-processing}), a bidirectional bounding-box overlap
criterion ($\geq 50\%$ of either the word or the target region), and a relaxed
span grouping filter that only excludes $[0,0,0,0]$ bounding boxes (previously,
the filter $x_1 < x_2 \wedge y_1 < y_2$ rejected $\sim$50\% of valid tokens
whose LayoutLMv3-normalised coordinates had $x_1 = x_2$), the V4 pipeline
achieves 79\% word-level and 70\% span-level link label recovery on the test
set (Section~\ref{sec:diagnostic-results}).

\subsection{Checkpointing}
\label{sec:checkpointing}

Model checkpoints are stored after each epoch and the best checkpoint (highest
validation combined F1) is retained. For Stage~4b, checkpoints include both
the encoder weights and the linker module weights as a single \texttt{.pt}
file. This enables a clean separation between training, prediction, and
evaluation, which is important for SLURM-based execution and reproducibility.

% ---------------------------------------------------------------------------
\section{Evaluation Protocol}
\label{sec:evaluation-protocol}

We implement a dual-threshold evaluation protocol:

\subsection{Text Matching: NED}
\label{sec:ned}

Normalised Edit Distance (NED) measures text accuracy:
\[
  \text{NED}(p, t) = \frac{\text{Levenshtein}(p, t)}{\max(|p|, |t|)}
  \tag{5.1}
\]
where $p$ is predicted text, $t$ is target text. Threshold: $\text{NED} < 0.2$.

\subsection{Location Matching: IoU}
\label{sec:iou}

Intersection over Union (IoU) measures spatial accuracy:
\[
  \text{IoU}(B_p, B_t) = \frac{\text{Area}(B_p \cap B_t)}{\text{Area}(B_p \cup B_t)}
  \tag{5.2}
\]
where $B_p$ is predicted box, $B_t$ is target box. Threshold: $\text{IoU} > 0.3$.

\subsection{Overall Metric: F1 Score}
\label{sec:f1}

A prediction matches ground truth if both $\text{NED} < 0.2$ and
$\text{IoU} > 0.3$.
\[
  \text{Precision} = \frac{\text{TP}}{\text{TP} + \text{FP}}, \quad
  \text{Recall} = \frac{\text{TP}}{\text{TP} + \text{FN}}, \quad
  F_1 = \frac{2 \cdot \text{Precision} \cdot \text{Recall}}{\text{Precision} + \text{Recall}}
  \tag{5.3}
\]

% ---------------------------------------------------------------------------
\section{Code Organisation}
\label{sec:code-organisation}

Our codebase is organised as a Python package:
\begin{verbatim}
Script/
|-- config.py          # Configuration parameters
|-- data_loader.py     # Dataset loading
|-- features.py        # Layout feature extraction
|-- visualization.py   # Plotting utilities
|-- metrics.py         # NED, IoU, F1 evaluation
|-- model.py           # LayoutLMv3KVPModel + BiaffineLinker
|-- train.py           # Stage 4 training loop
|-- main.py            # Pipeline orchestration
'-- utils.py           # Helper functions
\end{verbatim}

% ---------------------------------------------------------------------------
\section{Reproducibility}
\label{sec:reproducibility}

To ensure reproducibility:
\begin{itemize}
  \item Fixed random seed: \texttt{RANDOM\_SEED = 42}
  \item Version-controlled dependencies (\texttt{requirements.txt})
  \item Detailed configuration logging at run start
  \item Checkpoint saving at each epoch with best-model tracking
  \item Offline model caching for compute-node execution
  \item Separate login-node preparation and GPU-node training workflow on TinyGPU
\end{itemize}
