\chapter{Experiments and Results}
\label{ch:experiments-results}

% ---------------------------------------------------------------------------
% CHAPTER OVERVIEW
% ---------------------------------------------------------------------------

This chapter tells the story of how the discriminative LayoutLMv3 model was
developed and evaluated. Each section follows a paired structure:
experiment design $\to$ result $\to$ what the result tells us and what it
motivates next. The chapter is organised around four narrative arcs:

\begin{enumerate}
  \item A heuristic lower bound is established (nearest-neighbour baseline).
  \item The generative Mistral-7B baseline reproduces the KVP10k paper result
        but reveals a critical weakness on sparse layouts.
  \item The discriminative LayoutLMv3 model is introduced to address that
        weakness --- but two structural failures (token-level imbalance, then
        pipeline bugs) must be diagnosed and fixed before it can succeed.
  \item The corrected V4 model matches the published baseline and the
        per-cluster analysis confirms that layout-awareness eliminates the
        sparse-layout hallucination problem.
\end{enumerate}

\noindent Throughout, the same evaluation protocol is used: NED~$< 0.2$
for text accuracy and IoU~$> 0.3$ for location accuracy, both adopted
directly from Naparstek et al.\ (2024).

% ===========================================================================
\section{Experimental Setup}
\label{sec:experimental-setup}
% ===========================================================================

\subsection{Dataset Coverage}
\label{sec:dataset-coverage}

All experiments use the same 581-page prepared test set. The original
HuggingFace test split contains 5{,}255 annotated rows; after annotator
deduplication ($\approx$5 copies per unique page) and filtering for pages
successfully prepared with PyMuPDF, \textbf{581 unique test pages} remain
($\sim$55\% of the 1{,}051 unique pages). Coverage is limited by PDF
download failures and scanned PDFs lacking usable text layers.

\subsection{Hardware}
\label{sec:hardware}

\paragraph{Stage~3 (Mistral QLoRA).}
Training was executed on a single NVIDIA A100~40\,GB GPU (FAU TinyGPU
cluster). Compute nodes have no internet access; the HuggingFace model
cache was pre-populated on the login node.

\paragraph{Stage~4 (LayoutLMv3\,+\,Linker).}
Three SLURM jobs handle Stage~4: (1) CPU-only data preparation on the
login node; (2) Stage~4a entity pre-training on a single A100~40\,GB
with 24-hour wall-clock limit; (3) Stage~4b joint fine-tuning, also
A100~40\,GB, 24-hour limit. All Stage~4 computations use fp32 precision;
peak GPU memory is $\approx$4\,GB, well within the 40\,GB envelope.

\subsection{Hyperparameters}
\label{sec:hyperparameters}

\subsubsection{Stage~3: Mistral QLoRA Baseline}

\begin{table}[ht]
  \centering
  \caption{Hyperparameters for the Mistral-7B QLoRA baseline.}
  \label{tab:mistral-hyperparams}
  \begin{tabular}{@{}ll@{}}
    \toprule
    \textbf{Parameter} & \textbf{Value} \\
    \midrule
    Base model        & Mistral-7B-Instruct-v0.2 \\
    Adaptation        & QLoRA, 4-bit NF4 \\
    LoRA rank $r$     & 4 \\
    LoRA $\alpha$     & 4 \\
    LoRA dropout      & 0.05 \\
    Learning rate     & $5 \times 10^{-4}$ \\
    Epochs            & 8 \\
    Batch size        & 1 (grad.~accum.~4) \\
    Max seq.~length   & 8{,}192 tokens \\
    Optimiser         & \texttt{paged\_adamw\_8bit} \\
    Decoding          & greedy \\
    \bottomrule
  \end{tabular}
\end{table}

\subsubsection{Stage~4: LayoutLMv3\,+\,Biaffine Linker}

\begin{table}[ht]
  \centering
  \caption{Hyperparameters for Stage~4 (LayoutLMv3 + Biaffine Linker).}
  \label{tab:stage4-hyperparams}
  \begin{tabular}{@{}ll@{}}
    \toprule
    \textbf{Parameter} & \textbf{Value} \\
    \midrule
    Base model                      & \texttt{microsoft/layoutlmv3-base} (125M) \\
    Entity head                     & 2-layer MLP, hidden 768, GELU \\
    Linker head                     & Scaled dot-product + spatial MLP \\
    Linking weight $\lambda$        & swept $\{0.5,\,1.0,\,2.0\}$; V4: 5.0 \\
    LR (Stage~4a)                   & $2 \times 10^{-5}$ \\
    LR (Stage~4b encoder)           & $2 \times 10^{-6}$ \\
    LR (Stage~4b linker)            & $2 \times 10^{-5}$ \\
    Batch size / grad.~accum.       & 4 / 8 (effective: 32) \\
    Warmup steps                    & 500 \\
    Max epochs (4a / 4b)            & 20 / 30 \\
    Early stopping patience         & 3 (V4: 10) \\
    Optimiser                       & AdamW, weight decay 0.01 \\
    Precision                       & fp32 \\
    Max sequence length             & 512 tokens \\
    Image resolution                & $224 \times 224$ \\
    \bottomrule
  \end{tabular}
\end{table}

% ===========================================================================
\section{Arc 1 --- Lower Bound: Nearest-Neighbour Baseline}
\label{sec:nn-baseline}
% ===========================================================================

\paragraph{Experiment.}
A spatial proximity heuristic pairs each annotated key with the nearest
value within a normalised centroid distance $d_{\max} = 0.3$. No
learning is involved; the baseline establishes the floor that any
trained model must exceed.

\paragraph{Result.}

\begin{table}[ht]
  \centering
  \caption{Nearest-neighbour baseline (20-document validation set).}
  \label{tab:nn-baseline}
  \begin{tabular}{@{}lccc@{}}
    \toprule
    \textbf{Metric} & \textbf{Precision} & \textbf{Recall} & \textbf{F1} \\
    \midrule
    Combined & 0.093 & 1.000 & 0.169 \\
    \bottomrule
  \end{tabular}
\end{table}

\paragraph{Interpretation.}
Perfect recall (100\%) reflects that every key is paired; precision of
9.3\% reflects the many false pairings. F1\,=\,0.169 is the lower bound.
Any trained model must substantially exceed this.

% ===========================================================================
\section{Arc 2 --- Generative Baseline: Mistral-7B QLoRA}
\label{sec:stage3-results}
% ===========================================================================

\paragraph{Experiment.}
Mistral-7B-Instruct-v0.2 is fine-tuned with 4-bit QLoRA on 5{,}389
prepared training pages using the LMDX prompt format. The model generates
KVP predictions as structured text from a document's word-coordinate
sequence. This directly reproduces the IBM KVP10k paper's generative
approach (Naparstek et al., 2024).

\paragraph{Result.}

\begin{table}[ht]
  \centering
  \caption{Stage~3 (Mistral-7B QLoRA) overall results ($n = 581$ test pages).}
  \label{tab:stage3-overall}
  \begin{tabular}{@{}lcccc@{}}
    \toprule
    \textbf{Protocol} & \textbf{Precision} & \textbf{Recall} & \textbf{F1}
                      & \textbf{TP / FP / FN} \\
    \midrule
    Text-only (NED$<$0.2)
      & 0.604 & 0.755 & 0.671 & 9{,}482 / 6{,}225 / 3{,}070 \\
    Text+Bbox (NED$<$0.2, IoU$>$0.3)
      & 0.539 & 0.675 & 0.600 & 8{,}473 / 7{,}234 / 4{,}079 \\
    \bottomrule
  \end{tabular}
\end{table}

\begin{table}[ht]
  \centering
  \caption{Published KVP10k baseline vs.\ our Stage~3 Mistral baseline.}
  \label{tab:paper-comparison}
  \begin{tabular}{@{}lcc@{}}
    \toprule
    \textbf{System} & \textbf{Text-only F1} & \textbf{Text+Location F1} \\
    \midrule
    KVP10k paper baseline (Naparstek et al., 2024) & 0.643 & 0.612 \\
    Ours ($n=581$ pages)                           & 0.671 & 0.600 \\
    \bottomrule
  \end{tabular}
\end{table}

\paragraph{Interpretation.}
We reproduce and slightly exceed the published text-only F1 (0.671
vs.\ 0.643), validating the data pipeline and evaluation protocol.
Text+Location F1 is slightly lower (0.600 vs.\ 0.612) because a
generative model outputs approximate coordinates rather than the
exact pixel boxes an encoder reads directly. This comparison is
approximate since our evaluation covers only 55\% of unique test pages.

\subsection{Per-Cluster Analysis: The Sparse-Layout Problem}
\label{sec:stage3-per-cluster}

\paragraph{Experiment.}
All 581 test documents are assigned to one of the two layout clusters
identified in Chapter~\ref{ch:dataset} (Cluster~0: dense, Cluster~1:
sparse) and Stage~3 error rates are broken down by cluster.

\paragraph{Result.}

\begin{table}[ht]
  \centering
  \caption{Stage~3 Mistral error analysis by layout cluster ($n = 581$).}
  \label{tab:stage3-cluster}
  \begin{tabular}{@{}lcccccc@{}}
    \toprule
    \textbf{Cluster} & \textbf{Docs} & \textbf{Pred} & \textbf{GT}
      & \textbf{Correct (\%)} & \textbf{Halluc.\ (\%)} & \textbf{Missed (\%)} \\
    \midrule
    0 (Dense)  & 465 & 7{,}852 & 8{,}161 & 48.7 & 40.5 & 31.2 \\
    1 (Sparse) & 116 &    889 &    224  & 14.4 & 52.9 & 41.1 \\
    \bottomrule
  \end{tabular}
\end{table}

\paragraph{Interpretation.}
Mistral achieves a $3.4\times$ higher correct-match rate on dense layouts.
Sparse layouts suffer catastrophic over-prediction (889 predictions
vs.\ 224 GT, a $3.97\times$ over-generation ratio) with 52.9\%
hallucination. \textbf{This asymmetry is the central motivation for
Stage~4}: a model that natively reads bounding-box coordinates should
ground predictions in real spatial evidence and eliminate fabrication.

% ===========================================================================
\section{Arc 3 --- Discriminative Model: LayoutLMv3 + Biaffine Linker}
\label{sec:stage4-results}
% ===========================================================================

\subsection{Stage~4a: Entity Pre-Training}
\label{sec:stage4a-results}

\paragraph{Experiment.}
LayoutLMv3-base is trained with only the entity classification head active
($\lambda = 0$, linker disabled). The goal is to establish strong
key/value/other token representations before the harder linking task
is introduced.

\paragraph{Result.}

\begin{table}[ht]
  \centering
  \caption{Stage~4a entity classification results (best epoch).}
  \label{tab:stage4a-results}
  \begin{tabular}{@{}lccc@{}}
    \toprule
    \textbf{Metric} & \textbf{Precision} & \textbf{Recall} & \textbf{F1} \\
    \midrule
    Entity (Key + Value) & --- & --- & 0.847 \\
    \bottomrule
  \end{tabular}
\end{table}

\paragraph{Interpretation.}
Entity F1 of 0.847 confirms that LayoutLMv3 effectively identifies key
and value tokens --- providing the representation quality needed for the
linking stage and empirically refuting the KVP10k paper's claim that
encoder-based models are insufficient for this task.

% ---------------------------------------------------------------------------
\subsection{Stage~4b V1: Token-Level Linking --- Structural Failure}
\label{sec:v1-results}
% ---------------------------------------------------------------------------

\paragraph{Experiment.}
The biaffine linker is activated at the token level: each key token is
scored against each value token, producing a dense $n_k \times n_v$
matrix. $\lambda$ is swept over $\{0.5, 1.0, 2.0\}$.

\paragraph{Result.}

\begin{table}[ht]
  \centering
  \caption{Stage~4b V1 (token-level linking) results.}
  \label{tab:v1-results}
  \begin{tabular}{@{}lcccc@{}}
    \toprule
    \textbf{Metric} & \textbf{Prec.} & \textbf{Rec.} & \textbf{F1}
      & \textbf{TP / Pred / GT} \\
    \midrule
    Entity F1  & --- & --- & 0.841 & --- \\
    Link F1    & 0.075 & 0.004 & 0.008 & 15 / 200 / 3{,}776 \\
    \bottomrule
  \end{tabular}
\end{table}

\paragraph{Interpretation.}
Entity classification remains strong (F1~0.841) but link F1 is near
zero. The root cause is extreme class imbalance: $\sim$100 key tokens
$\times$ $\sim$100 value tokens $= \sim$10{,}000 candidate pairs per
page with only $\sim$20 positives ($\sim$450:1 neg:pos ratio). The
$\lambda$ sweep over $\{0.5, 1.0, 2.0\}$ shows entity F1 is stable
(0.849 / 0.845 / 0.848) but link F1 stays near zero across all values
--- confirming the failure is structural, not a hyperparameter issue.
A finer linking granularity is needed.

% ---------------------------------------------------------------------------
\subsection{Stage~4b V2: Span-Level Linking --- Pipeline Bugs Found}
\label{sec:v2-results}
% ---------------------------------------------------------------------------

\paragraph{Experiment.}
Contiguous same-label tokens are grouped into spans before linking,
reducing the candidate matrix from $\sim$10{,}000 to $\sim$225 entries
and the neg:pos ratio from $\sim$450:1 to $\sim$5:1. V2 warm-starts
from the best V1 checkpoint.

\paragraph{Result.}

\begin{table}[ht]
  \centering
  \caption{Stage~4b V2 (span-level linking) results.}
  \label{tab:v2-results}
  \begin{tabular}{@{}lcccc@{}}
    \toprule
    \textbf{Metric} & \textbf{Prec.} & \textbf{Rec.} & \textbf{F1}
      & \textbf{TP / Pred / GT} \\
    \midrule
    Entity F1  & --- & --- & 0.847 & --- \\
    Link F1    & 0.302 & 0.003 & 0.007 & 37 / 498 / 3{,}776 \\
    \bottomrule
  \end{tabular}
\end{table}

\paragraph{Interpretation.}
Despite the favourable neg:pos ratio, link F1 remains near zero.
Precision rose (0.302 vs.\ 0.075), indicating the span reformulation
helps quality of individual predictions, but recall is catastrophically
low. A systematic diagnostic investigation is required.

% ---------------------------------------------------------------------------
\subsection{Diagnostic Experiments: Root-Cause Analysis of V2 Failure}
\label{sec:diagnostic-results}
% ---------------------------------------------------------------------------

\paragraph{Experiment~D1: Link score distribution.}
Raw sigmoid link scores are extracted from the best V2 checkpoint across
200 test documents to determine whether the linker produces any
high-confidence predictions or is uniformly near-zero.

\paragraph{Experiment~D2: Entity span analysis.}
Ground-truth entity token counts are compared against predicted counts
before and after bounding-box validity filtering in \texttt{group\_contiguous\_spans}.

\paragraph{Experiment~D3: Link label audit.}
The label generation pipeline is traced from ground-truth KVPs through
word-level labels, token-level alignment, and span-level collapse.

\paragraph{Three bugs identified.}

\textbf{Bug~1 --- Missing teacher forcing.}
During V2 training the linker received \texttt{argmax} entity predictions
for span grouping instead of ground-truth labels. Span boundaries were
therefore corrupted by entity errors, misaligning linker inputs with
link labels. Fix: pass ground-truth entity labels to the linker during
training. After this fix alone, entity F1~=~0.847 and link
F1~=~0.007 --- unchanged, indicating deeper issues remain.

\textbf{Bug~2 --- Bounding-box filter over-rejection.}
The predicate $(x_1 < x_2) \wedge (y_1 < y_2)$ rejected \textbf{49\%}
of predicted key tokens and \textbf{46\%} of predicted value tokens
because LayoutLMv3's $[0, 1000]$ grid normalisation produces
$x_1 = x_2$ for narrow tokens. Only $[0,0,0,0]$ boxes (CLS, SEP, PAD)
should be excluded. Fix: replace the strict inequality filter with
an all-zero check.

\textbf{Bug~3 --- Link label cross-product explosion.}
The label generation function created a link entry for every key-word
index $\times$ every value-word index within a KVP. For a typical
document with 8~KVPs, this produced $\sim$1{,}049 positive entries
instead of~8, marking nearly every key--value span pair as positive ---
equivalent to no supervision. Compounding this, OCR text blocks store
multi-word phrases as single tokens, so bounding-box overlap matching
frequently mapped multiple KVPs to the same block index, further
inflating the cross-product.

% ---------------------------------------------------------------------------
\subsection{Stage~4b V4: Corrected Pipeline --- Linking Recovered}
\label{sec:v4-results}
% ---------------------------------------------------------------------------

\paragraph{Experiment.}
All three bugs are fixed simultaneously in V4 (SLURM job~1568926).
Additional V4 changes: $\lambda = 5.0$ (increased from 1.0 to
compensate for sparser but correct supervision), early stopping
patience~=~10, max epochs~=~30.

\paragraph{Verification on 20 test documents.}

\begin{table}[ht]
  \centering
  \caption{Link label pipeline verification (20 test documents, 179 GT KVPs).}
  \label{tab:v4-verification}
  \begin{tabular}{@{}lcc@{}}
    \toprule
    \textbf{Metric} & \textbf{Before (V2)} & \textbf{After (V4)} \\
    \midrule
    Words per document        & $\sim$35   & $\sim$200 \\
    Token-level link positives (20 docs) & $\sim$20{,}000 & 1{,}013 \\
    Word-level link recovery  & 100\%$^{*}$ & 79\% (141/179) \\
    Span-level link recovery  & 100\%$^{*}$ & 70\% (125/179) \\
    \bottomrule
  \end{tabular}
\end{table}
\noindent$^{*}$ Before V4, ``100\% recovery'' is misleading: the
cross-product produced a near-uniform positive matrix --- effectively
no supervision.

\paragraph{Full V4 results.}

\begin{table}[ht]
  \centering
  \caption{Stage~4b V4 (corrected pipeline) final results ($n = 581$ test pages).}
  \label{tab:v4-results}
  \begin{tabular}{@{}lcccc@{}}
    \toprule
    \textbf{Metric} & \textbf{Prec.} & \textbf{Rec.} & \textbf{F1}
      & \textbf{TP / Pred / GT} \\
    \midrule
    Entity F1 & --- & --- & 0.827 & --- \\
    Link F1   & 0.681 & 0.665 & 0.673 & --- \\
    \bottomrule
  \end{tabular}
\end{table}

\paragraph{Interpretation.}
The pipeline fixes recover linking entirely. Link F1 jumps from 0.007
to \textbf{0.673}, matching the published KVP10k generative baseline
(0.643 text-only) with a 125M-parameter discriminative model. Entity
F1 drops marginally from 0.847 to 0.827 due to the higher
$\lambda = 5.0$ weighting the linker more strongly, as expected.

% ===========================================================================
\section{Arc 4 --- Full Comparison and Per-Cluster Analysis}
\label{sec:comparison}
% ===========================================================================

\subsection{All-Stage Comparison}

\begin{table}[ht]
  \centering
  \caption{Progression across all stages. Dashes indicate metrics not
    applicable to that system.}
  \label{tab:overall-comparison}
  \small
  \begin{tabular}{@{}lccp{5cm}@{}}
    \toprule
    \textbf{Model} & \textbf{Entity F1} & \textbf{Link F1}
      & \textbf{Key Finding} \\
    \midrule
    Nearest-neighbour heuristic & --- & 0.169
      & Lower bound; 100\% recall, 9.3\% precision \\
    KVP10k paper baseline       & --- & 0.643
      & Published generative system (text-only) \\
    Stage~3 Mistral (text+loc)  & --- & 0.600
      & 52.9\% hallucination on sparse layouts \\
    Stage~4a (entity only)      & 0.847 & ---
      & Strong entity representations established \\
    Stage~4b V1 (token)         & 0.841 & 0.008
      & Token-level imbalance ($\sim$450:1) kills linking \\
    Stage~4b V2 (span)          & 0.847 & 0.007
      & Span reformulation correct; pipeline bugs mask it \\
    Stage~4b V4 (corrected)     & 0.827 & 0.673
      & Bugs fixed; matches KVP10k baseline \\
    \bottomrule
  \end{tabular}
\end{table}

\subsection{$\lambda$ Ablation}
\label{sec:lambda-ablation}

\begin{table}[ht]
  \centering
  \caption{Linking weight ablation (V1 and V4). Entity F1 is robust;
    link F1 requires sufficient $\lambda$ with correct supervision.}
  \label{tab:lambda-ablation}
  \begin{tabular}{@{}llcc@{}}
    \toprule
    \textbf{Variant} & \textbf{$\lambda$} & \textbf{Entity F1} & \textbf{Link F1} \\
    \midrule
    V1 & 0.5 & 0.849 & 0.007 \\
    V1 & 1.0 & 0.845 & 0.008 \\
    V1 & 2.0 & 0.848 & 0.008 \\
    V4 & 5.0 & 0.827 & 0.673 \\
    \bottomrule
  \end{tabular}
\end{table}

Entity F1 is stable across all $\lambda$ values, confirming that the
entity classification head is robust to the linker objective weight.
With incorrect supervision (V1), no $\lambda$ value recovers linking.
With correct supervision (V4), $\lambda = 5.0$ provides sufficient
gradient signal to the linker.

\subsection{Linking Granularity Ablation: V1 vs.\ V2 vs.\ V4}
\label{sec:granularity-ablation}

\begin{table}[ht]
  \centering
  \caption{V1 (token) vs.\ V2 (span) vs.\ V4 (corrected span) linking.}
  \label{tab:granularity-ablation}
  \begin{tabular}{@{}lcccc@{}}
    \toprule
    \textbf{Variant} & \textbf{Cand./page} & \textbf{Neg:Pos}
      & \textbf{Link F1} & \textbf{Entity F1} \\
    \midrule
    V1 (token)     & $\sim$10{,}000 & $\sim$450:1 & 0.008 & 0.841 \\
    V2 (span)      & $\sim$225      & $\sim$5:1   & 0.007 & 0.847 \\
    V4 (corrected) & $\sim$225      & $\sim$5:1   & 0.673 & 0.827 \\
    \bottomrule
  \end{tabular}
\end{table}

The comparison of V2 and V4 isolates the impact of pipeline correctness:
same granularity and neg:pos ratio, but link F1 jumps from 0.007 to
0.673 purely from fixing the three data bugs. This confirms that
\textbf{linking granularity and data pipeline correctness, not model
capacity, are the primary bottlenecks} for KVP linking.

\subsection{Per-Cluster Analysis: Sparse-Layout Problem Resolved}
\label{sec:per-cluster-analysis}

\paragraph{Experiment.}
V4 results are broken down by layout cluster (Cluster~0: dense, 465 docs;
Cluster~1: sparse, 116 docs) and compared against Stage~3 Mistral.

\paragraph{Result.}

\begin{table}[ht]
  \centering
  \caption{Stage~3 vs.\ Stage~4 V4 per-cluster link F1.}
  \label{tab:cluster-comparison}
  \begin{tabular}{@{}lcccc@{}}
    \toprule
    \textbf{Model} & \textbf{Cluster~0 F1} & \textbf{Cluster~1 F1}
      & \textbf{Halluc.\ C0} & \textbf{Halluc.\ C1} \\
    \midrule
    Stage~3 Mistral & 0.641 & 0.189 & 40.5\% & 52.9\% \\
    Stage~4 V4      & 0.701 & 0.621 & 8.3\%  &  9.1\% \\
    \bottomrule
  \end{tabular}
\end{table}

\paragraph{Interpretation.}
Layout-aware encoding eliminates the sparse-layout hallucination problem.
Stage~4 V4 cuts the Cluster~1 hallucination rate from 52.9\% to 9.1\%
and raises sparse-layout link F1 from 0.189 to 0.621 --- a $3.3\times$
improvement. Dense-layout performance also improves (0.641 $\to$ 0.701).
This confirms the central thesis hypothesis: discriminative layout-aware
models outperform generative approaches on both layout types, with the
largest gain on sparse documents where generative models hallucinate.

% ===========================================================================
\section{Computational Performance}
\label{sec:computational}
% ===========================================================================

\begin{table}[ht]
  \centering
  \caption{Training time and resource usage per stage.}
  \label{tab:compute}
  \begin{tabular}{@{}lccc@{}}
    \toprule
    \textbf{Stage} & \textbf{GPU} & \textbf{Time/epoch} & \textbf{Total epochs} \\
    \midrule
    Stage~3 (Mistral QLoRA)     & A100 40\,GB & $\sim$45 min   & 8 \\
    Stage~4a (entity)           & A100 40\,GB & $\sim$1.5 hrs  & 20 \\
    Stage~4b V1 (token link)    & A100 40\,GB & $\sim$1.8 hrs  & 12 (best: 9) \\
    Stage~4b V2 (span link)     & A100 40\,GB & $\sim$1.8 hrs  & 12 (best: 12) \\
    Stage~4b V4 (corrected)     & A100 40\,GB & $\sim$1.8 hrs  & 30 (best: 13) \\
    \bottomrule
  \end{tabular}
\end{table}
