\chapter{Discussion}
\label{ch:discussion}

% Renamed from 08_discussion.tex to 07_discussion.tex to match new chapter order.
% Content preserved exactly from original 08_discussion.tex.

\section{Interpretation of Results}
\label{sec:interpretation}

The results across all stages tell a coherent story about the interplay
between model architecture and data pipeline quality in document
understanding tasks.

The generative Mistral-7B baseline (Stage~3) achieves competitive
text-only F1 (0.671) but reveals a fundamental weakness: it hallucinates
key-value pairs at a 52.9\% rate on sparse-layout documents (Cluster~1),
where the absence of dense spatial cues leads the model to fabricate
plausible-looking but incorrect KVP structures. This is a well-documented
limitation of generative models in structured prediction tasks
\cite{maynez2020faithfulness}: they optimise for fluency over faithfulness,
producing outputs that are linguistically coherent but factually grounded
only in the training distribution rather than the specific input document.

The discriminative LayoutLMv3 approach (Stage~4) addresses this by
construction: it can only classify tokens that are actually present in the
document, making hallucination structurally impossible. The near-zero link
F1 of V1 and V2 initially appeared to be a model capacity problem, but
systematic diagnostic analysis revealed that the root cause was a
data-pipeline defect: corrupted link supervision (cross-product explosion)
and aggressive bounding-box filtering that discarded nearly half of valid
entity tokens. After correcting these bugs, V4 achieves link F1~=~0.673,
matching the published KVP10k generative baseline.

\section{Generative vs.\ Discriminative Trade-offs}
\label{sec:gen-vs-disc}

The comparison between Stage~3 and Stage~4 V4 illuminates the trade-offs
between generative and discriminative approaches for structured document
extraction:

\begin{itemize}
  \item \textbf{Hallucination:} Generative models hallucinate on sparse
  layouts (52.9\% Cluster~1); discriminative models do not (9.1\%).
  \item \textbf{Scale:} Mistral-7B uses $\sim$56$\times$ more parameters
  than LayoutLMv3-base (7B vs.\ 125M) to achieve comparable overall F1.
  \item \textbf{Spatial grounding:} LayoutLMv3's native 2D position
  embeddings provide richer spatial context than the coordinate-string
  encoding used in the LMDX prompt format.
  \item \textbf{Generalisation:} Discriminative models require task-specific
  linking heads; generative models produce structured output directly
  but are more sensitive to prompt formatting and output parsing.
  \item \textbf{Data sensitivity:} Stage~4 is significantly more sensitive
  to data pipeline correctness than Stage~3 --- a design consideration
  for practitioners deploying discriminative document models.
\end{itemize}

\section{Limitations}
\label{sec:limitations}

\subsection{Dataset Coverage}

All experiments are conducted on 55\% of the KVP10k test set (581 of
1{,}051 unique pages) due to PDF download failures and scanned PDFs
lacking usable text layers. This coverage limitation means comparisons
with the published KVP10k baseline are approximate: we evaluate on a
different (and potentially non-representative) subset. The missing
45\% of pages is disproportionately composed of scanned documents,
which may be harder than the average page, suggesting our reported
performance may be optimistically biased.

\subsection{OCR Dependency}

The entire Stage~3 and Stage~4 pipeline depends on high-quality text
layers in the source PDFs. Documents without embedded text (scanned
images) are excluded. An OCR-free approach (e.g., Pix2Struct, Donut)
would eliminate this dependency and extend coverage to the full dataset.

\subsection{Single Dataset}

All results are reported on KVP10k only. The extent to which the
LayoutLMv3 + Biaffine Linker architecture generalises to other
document understanding benchmarks (e.g., FUNSD, CORD, SROIE) remains
an open question.

\subsection{Fixed Hyperparameters}

Due to HPC resource constraints, the $\lambda$ sweep is limited to
three values for V1 and a single value for V4. A more exhaustive
hyperparameter search (including learning rate, warmup schedule,
and entity/linker head architecture) might yield further improvements.

\section{Embedding Analysis}
\label{sec:embedding-analysis}

To gain qualitative insight into the learned representations,
t-SNE projections of the LayoutLMv3 encoder's final-layer hidden
states are computed for 50 randomly sampled test documents.
Tokens are coloured by their ground-truth label (Key, Value, Other).

The t-SNE visualisation reveals three loosely separated clusters
aligned with Key, Value, and Other labels, with substantial overlap
between Key and Value tokens. This is consistent with the high entity
F1 (0.827): the model has learnt layout-dependent representations
that distinguish roles, but the embedding geometry does not fully
separate the classes --- motivating the use of a learned linking
head rather than relying on embedding distance alone.

\section{Future Work}
\label{sec:future-work}

\begin{enumerate}
  \item \textbf{OCR-free extension:} Replace the PyMuPDF text-layer
  pipeline with an end-to-end vision model (Donut, Pix2Struct) to
  handle scanned documents and extend KVP10k coverage to 100\%.
  \item \textbf{Cross-dataset generalisation:} Evaluate the V4
  architecture on FUNSD, CORD, and SROIE to assess transfer without
  retraining.
  \item \textbf{Larger encoders:} Evaluate LayoutLMv3-large (368M)
  and LayoutXLM to assess whether additional capacity improves the
  linking head given correct supervision.
  \item \textbf{Active learning:} The cluster analysis shows that
  sparse-layout documents are harder; an active-learning loop that
  selects informative sparse examples for annotation could improve
  Cluster~1 performance further.
  \item \textbf{Robustness evaluation:} Assess model stability under
  document rotation ($0$--$15^\circ$) and occlusion (masking
  10--30\% of page area).
\end{enumerate}
