\chapter{Conclusion}
\label{ch:conclusion}

% Renamed from 09_conclusion.tex to 08_conclusion.tex to match new chapter order.
% Content preserved exactly from original 09_conclusion.tex.

\section{Summary of Contributions}
\label{sec:contributions}

This thesis has made the following contributions to the problem of
automatic key-value pair extraction from business documents:

\begin{enumerate}
  \item \textbf{Empirical refutation of the KVP10k paper's claim.}
  The KVP10k paper (Naparstek et al., 2024) argued, without direct
  empirical evidence, that encoder-based models are insufficient for
  KVP linking. This thesis demonstrates that LayoutLMv3 with a learned
  Biaffine Linker achieves link F1~=~0.673 --- matching the published
  generative baseline --- with 56$\times$ fewer parameters.

  \item \textbf{Two-phase training strategy.}
  A Stage~4a entity pre-training phase followed by Stage~4b joint
  fine-tuning stabilises training and achieves entity F1~=~0.847,
  providing strong representations before the harder linking task
  is introduced.

  \item \textbf{Granularity analysis of KVP linking.}
  Token-level linking (V1) fails due to $\sim$450:1 class imbalance.
  Span-level linking (V2/V4) reduces this to $\sim$5:1 and, with
  correct supervision, enables effective linking. \textbf{Linking
  granularity, not model capacity, is the primary bottleneck.}

  \item \textbf{Systematic diagnostic methodology.}
  Three compounding data-pipeline bugs (missing teacher forcing,
  over-aggressive bounding-box filtering, link label cross-product
  explosion) were identified through structured diagnostic experiments
  and corrected in V4. This diagnostic methodology is transferable to
  other document understanding pipelines.

  \item \textbf{Layout-cluster analysis.}
  K-means clustering identifies two document layout types (dense:
  80\%, sparse: 20\%). Stage~4 V4 cuts sparse-layout hallucination
  from 52.9\% (Mistral) to 9.1\% and raises Cluster~1 link F1 from
  0.189 to 0.621, demonstrating that layout-aware discriminative
  models are particularly beneficial for sparse documents.
\end{enumerate}

\section{Final Remarks}
\label{sec:final-remarks}

The central finding of this thesis is that the failure of early
LayoutLMv3 linking experiments was not a model capacity problem but
a data pipeline problem. This has an important practical implication:
before attributing poor linking performance to architectural limitations,
practitioners should carefully audit their label generation pipeline ---
particularly the cross-product explosion pattern in multi-word KVP
annotations and the bounding-box validity filtering in span grouping.

The corrected V4 model demonstrates that a 125M-parameter discriminative
encoder with a lightweight linking head can match a 7B-parameter
generative model on KVP extraction while eliminating spatial hallucination.
This result supports the value of investing in data pipeline correctness
and layout-aware architectures for structured document understanding.
