\chapter{Discussion}
\label{ch:discussion}

\section{Interpretation of Results}
\label{sec:interpretation}

\subsection{Entity Classification Success}

The LayoutLMv3 V4 entity classifier achieves F1~$= 0.827$, demonstrating that
the multimodal encoder effectively distinguishes key, value, and other tokens
using joint text-layout-image representations. This result validates the
choice of a discriminative formulation over the generative baseline: by
constraining predictions to existing OCR tokens, the model eliminates
spatial hallucination entirely.

\subsection{Token-Level Linking Failure}

The near-zero link F1 ($0.008$) of V1 is the central negative result of this
thesis. The failure is attributable to the granularity mismatch between the
linking formulation (token-to-token) and the task structure (span-to-span).
With $\sim$450:1 negative-to-positive ratio at the token level, the BCE loss
is dominated by negative examples even with \texttt{pos\_weight} correction.
This finding aligns with SPADE~\cite{spade} and BROS~\cite{bros}, which both
operate at the segment/span level rather than the token level.

\subsection{Comparison to Baselines}

Compared to the Stage~3 Mistral baseline (text+location F1~$= 0.600$),
the discriminative LayoutLMv3 model provides strong entity extraction but
the initial V2 span-level linker fails to improve linking over V1 (both near
zero).  The published KVP10k baseline achieves linking F1~$= 0.659$ using
a text-only generative approach.  The diagnostic analysis
(Section~\ref{sec:diagnostic-results}) revealed that V2's failure was caused
not by the span-level architecture itself but by three compounding data
pipeline bugs.  Once all three pipeline bugs were corrected, the V4 model
achieves link F1~$= 0.673$ (threshold $= 0.5$) --- matching the KVP10k
paper's published baseline with a 125M-parameter discriminative model versus
their proprietary generative system.

\section{Root-Cause Analysis}
\label{sec:root-cause-analysis}

The progression from V1 to V4 illustrates a common but underreported pattern
in structured prediction pipelines: architectural improvements cannot
compensate for corrupted supervision.  Each bug individually would have been
sufficient to prevent learning:

\begin{enumerate}
  \item \textbf{Missing teacher forcing} (Bug~1): The linker was trained on
  spans derived from noisy \texttt{argmax} predictions rather than
  ground-truth labels.  Span boundaries were inconsistent with the
  ground-truth link labels, creating a train--test mismatch that prevented
  convergence.

  \item \textbf{Bounding box filter} (Bug~2): The strict
  $(x_1 < x_2) \wedge (y_1 < y_2)$ predicate rejected tokens whose
  normalised coordinates had $x_1 = x_2$ after LayoutLMv3's integer
  quantisation to $[0, 1000]$.  This removed 49\% of predicted key tokens
  and 46\% of value tokens --- halfing the number of entity spans available
  for linking.

  \item \textbf{Cross-product link labels} (Bug~3): The label generator
  created all-pairs links between key and value words rather than a single
  representative link per KVP.  Combined with OCR blocks that concatenated
  multiple words into single ``word'' entries, this produced $\sim$1{,}000
  positive link entries per document instead of $\sim$8, making the
  span-level collapse produce near-uniform positive labels.
\end{enumerate}

\noindent The critical insight is that Bug~3 was invisible at the token level
(V1), where the dense $n_k \times n_v$ matrix already had many positives.
It only became observable when the span-level collapse was introduced in V2,
because the small $K \times V$ matrix should have been sparse but was not.
This underscores the value of the span-level reformulation as a diagnostic
tool, not only an architectural improvement.

\section{Error Analysis}
\label{sec:error_analysis}

\subsection{Entity Classification Errors}

Entity errors fall into three categories: (i)~boundary errors, where partial
spans are predicted (e.g., missing the last token of a multi-word key);
(ii)~type confusion, where a key is predicted as a value or vice versa; and
(iii)~false positives on non-entity tokens that appear near key-value regions.

\subsection{Linking Errors (V1)}

V1 linking errors are dominated by false negatives: the model predicts only
200 links against 3{,}776 ground-truth links. Of the 200 predictions, 15
are correct (7.5\% precision). The model adopts an overly conservative
strategy, preferring to predict no link rather than risk a false positive ---
a rational response to the extreme class imbalance in the training signal.

\subsection{V2 Linking Errors}

V2 span-level linking achieves link F1\,=\,0.007, marginally below V1's
0.008.  The model predicts 498~links across 581 test documents, of which
37 are correct (7.4\% precision).  Recall is 0.3\%, with only 37 of
3{,}776 ground-truth links recovered.

The error pattern differs from V1 in a revealing way: the linker produces
more predictions (498~vs.\ 200) but no improvement in true positives.  The
additional predictions are false positives caused by the corrupted
link supervision (Bug~3), which trained the linker to associate nearly any
span pair.  With the V4 corrections, the linker receives correct sparse
supervision where each KVP produces exactly one positive span link.

A supplementary experiment---V3 (linker-only), where the encoder and
entity classifier were frozen and only the linker was trained---achieved
83\% link accuracy on the validation set, demonstrating that the
\texttt{SpanBiaffineLinker} architecture itself is capable of learning
correct associations when given clean input representations.  The test-set
link F1 of V3 (0.018) remained low because it inherited the same sparse
entity spans from the bbox-filtered entity classifier (Bug~2).

\section{Update to the State of the Art at Submission}
\label{sec:sota_update}

The literature review in Chapter~\ref{ch:literature} adopts a
time-bounded view of the state of the art, fixed at the start of the
project (February~2026). During the execution of this thesis, we identified
two additional models --- SPADE~\cite{spade} and BROS~\cite{bros} --- that
are directly relevant to our linking methodology. Both models were published
before our project start but were not included in the original literature
survey. They have been integrated into the literature review
(Section~\ref{subsec:entity_linking}) and directly informed the design of the
V2 span-level linker.

From an evaluation perspective, our experimental comparisons remain anchored
to the KVP10k baseline. SPADE and BROS are evaluated on CORD and FUNSD
respectively, not on KVP10k, so direct numerical comparison is not possible.
However, their architectural insight --- that entity linking should operate at
the segment/span level --- is validated by our V1 failure analysis.

\section{Limitations and Future Work}
\label{sec:limitations}

\subsection{Current Limitations}

\begin{itemize}
    \item \textbf{Training data coverage}: Our training set covers only
    55.8\% of unique KVP10k pages due to PDF download failures and
    scanned documents lacking text layers. An OCR-based pipeline (e.g.,
    Tesseract) would increase coverage.

    \item \textbf{LoRA rank}: Hardware constraints (40\,GB A100) forced
    the Mistral baseline to use LoRA rank $r = 4$ instead of the
    estimated $r \geq 16$ in the official baseline. This limits the
    fairness of the Stage~3 vs.\ paper comparison.

    \item \textbf{Single-page limitation}: The current pipeline processes
    one page at a time; multi-page documents with cross-page key-value
    pairs cannot be handled.
\end{itemize}

\subsection{Future Directions}
\begin{itemize}
    \item Multi-page document handling with cross-page linking
    \item End-to-end OCR integration to increase data coverage
    \item Cross-lingual evaluation on XFUND
    \item Contrastive pre-training for the linker head
\end{itemize}
