\chapter{Conclusion}
\label{ch:conclusion}

\section{Summary of Contributions}
\label{sec:summary}

This thesis presented a comprehensive study on \kvp{} extraction from business
documents using LayoutLMv3 with a learned span-level linker. The main
contributions are:

\begin{enumerate}
    \item \textbf{Dataset Analysis}: Characterised the KVP10k dataset through
    layout clustering, revealing two distinct layout types and strong spatial
    proximity patterns in \kvps{} (Chapter~\ref{ch:dataset}).

    \item \textbf{Generative Baseline}: Reproduced the KVP10k Mistral-7B
    baseline and identified its failure modes: hallucination on sparse layouts
    (52.9\% hallucination rate) and weak spatial grounding
    (Chapter~\ref{ch:results}).

    \item \textbf{Discriminative Architecture}: Implemented a LayoutLMv3-based
    entity classifier achieving F1~$= 0.827$, eliminating spatial
    hallucination through grounded token classification
    (Chapter~\ref{ch:methodology}).

    \item \textbf{Linking Granularity Analysis}: Demonstrated that token-level
    linking fails due to extreme class imbalance ($\sim$450:1), yielding link
    F1~$= 0.008$. Designed and implemented span-level linking (V2) to address
    this mismatch (Chapter~\ref{ch:methodology}).

    \item \textbf{Pipeline Debugging and V4 Results}: Diagnosed three
    compounding data-pipeline bugs (missing teacher forcing, over-strict
    bounding-box filter, cross-product link label explosion) that masked
    span-level linking performance. The corrected V4 model achieves
    link F1~$= 0.673$, on par with the published KVP10k baseline using
    a 125M-parameter discriminative model
    (Chapters~\ref{ch:results}--\ref{ch:discussion}).

    \item \textbf{Implementation}: Released a modular, reproducible codebase
    with CONFIG-controlled experiments.
\end{enumerate}

\section{Research Questions Revisited}
\label{sec:rq_revisited}

\subsection{RQ1: Layout Characteristics}
Layout clustering reveals two distinct document types (dense and sparse) with
82.7\% of variance explained by two principal components. \kvps{} exhibit
strong spatial proximity (mean distance = 0.10 normalised units), confirming
that spatial features are essential for accurate linking.

\subsection{RQ2: Discriminative vs.\ Generative}
The discriminative LayoutLMv3 model achieves entity F1~$= 0.827$ with no
spatial hallucination, compared to Mistral's text+location F1~$= 0.600$ with
significant hallucination. Layout-aware discriminative models outperform
text-only generative approaches on spatially-grounded extraction.

\subsection{RQ3: Linking Granularity}
Token-level linking (V1) fails decisively (F1~$= 0.008$) due to the
$\sim$450:1 class imbalance inherent in token-to-token scoring. Span-level
linking (V2) reduces this to $\sim$5:1 by grouping contiguous tokens before
scoring. However, three data-pipeline bugs initially masked this benefit,
producing link F1 of only 0.007. After systematic diagnosis and correction,
the V4 model achieves link F1~$= 0.673$, confirming that span-level linking
is effective when paired with correct supervision.

\subsection{RQ4: Success and Failure Factors}
Model performance depends strongly on document layout type. Dense layouts
benefit from rich contextual signals; sparse layouts challenge both generative
(hallucination) and discriminative (fewer training examples) approaches.
Linking granularity, not model capacity, is the primary bottleneck for
key-value association.

\section{Practical Implications}
\label{sec:implications}

This work demonstrates that discriminative multimodal models are well-suited
for enterprise document processing: they provide spatially grounded
predictions, eliminate hallucination, and operate efficiently on modest
hardware (single GPU, 125M parameters). The modular pipeline design enables
adaptation to new document types with minimal re-engineering.

\section{Closing Remarks}
\label{sec:closing}

This thesis demonstrates that effective \kvp{} extraction requires both
layout-aware encoding and appropriate linking granularity. While entity
classification is largely solved by LayoutLMv3 (F1~$= 0.827$), the linking
problem demands careful attention to data pipeline correctness, class
imbalance, and task structure. The progression from V1 through V4 illustrates
that architectural improvements --- such as span-level linking --- can only
realise their potential when the supervision signal is correct. The final
V4 model achieves link F1~$= 0.673$, matching the published KVP10k baseline
with a compact 125M-parameter discriminative model.
