% ==============================================================================
% STAGE 4 METHODOLOGY - READY TO ADD TO Chapter 3
% ==============================================================================
% This section is fully writeable without awaiting training results.
% It describes architectural design decisions, not empirical findings.
% 
% INSERT THIS INTO: LaTeX_Thesis/chapters/03_methodology.tex
% LOCATION: After Section 1 "Overview", before existing encoder/decoder sections
% ==============================================================================

\section{Stage 4: Discriminative Key-Value Pair Extraction}
\label{sec:stage4_methodology}

Stage 4 addresses the core KVP extraction task through a discriminative learning approach combining layout-aware encoding with explicit relation modeling. This section describes the architecture, design choices, and training procedure developed for Stages 4a (entity classification) and 4b (entity + link prediction).

\subsection{Architectural Overview}
\label{sec:stage4_architecture}

The Stage 4 architecture departs from the generative approach of Stage 3 by formulating KVP extraction as a structured prediction problem with two sub-tasks:

\begin{enumerate}
    \item \textbf{Entity Recognition} (Stages 4a and 4b): Classify each token in the document as \texttt{Other}, \texttt{Key}, or \texttt{Value}.
    \item \textbf{Relation Linking} (Stage 4b only): Predict which Value tokens correspond to each Key token via scored attention.
\end{enumerate}

This decomposition offers two distinct advantages over the generative baseline:

\begin{enumerate}
    \item \textbf{Spatial grounding}: Discriminative models inherit bounding box predictions from the encoder without hallucination, whereas generative models must construct bounding boxes from text representations.
    \item \textbf{Unsupervised domain adaptation}: Transition learning can be invoked from Stage 3 (text encoding) into Stage 4 (spatial linking), leveraging the lmdx\_text representations already available.
\end{enumerate}

The encoder backbone is LayoutLMv3 (microsoft/layoutlmv3-base), chosen for its state-of-the-art multimodal fusion and proven effectiveness on form-understanding benchmarks (FUNSD, CORD, DocVQA). Stage 4a adds a simple softmax classifier over the encoder output to produce entity labels; Stage 4b adds an additional scoring module to predict relation scores.

\subsection{Why LayoutLMv3 as Encoder}
\label{sec:layoutlmv3_justification}

LayoutLMv3 was selected over alternative layout-aware encoders (LayoutLMv2, VisualBERT, Donut) for the following reasons:

\begin{enumerate}
    \item \textbf{Unified modality fusion}: Unlike LayoutLMv2 (which appends a CNN), LayoutLMv3 uses a single transformer backbone for both text and image, simplifying integration with downstream classification heads and reducing parameter count.
    
    \item \textbf{Spatial awareness without overhead}: LayoutLMv3 preserves 2D position embeddings (bounding box coordinates normalized to $[0, 1000]$ range) within its transformer layers, enabling the model to implicitly learn spatial relationships without explicit graph construction.
    
    \item \textbf{Training stability}: Pre-training on masked language modeling, masked image modeling, and word-patch alignment objectives produces robust representations, reducing fine-tuning instability compared to models pre-trained on vision-only objectives.
    
    \item \textbf{Proven downstream performance}: LayoutLMv3 achieves state-of-the-art results on FUNSD (F1 = 0.96) and CORD (F1 = 0.97) at the time of this thesis (February 2026), outperforming LayoutLMv2 and other layout-aware competitors on form understanding benchmarks.
\end{enumerate}

\subsection{Stage 4a: Entity Classification}
\label{sec:stage4a_entity_classification}

\subsubsection{Architecture}

Stage 4a is the simplest and most data-efficient variant. The LayoutLMv3 encoder output (sequence of hidden states, one per token) is fed to a linear projection layer followed by softmax:

\[
P(\text{label}_i \mid \text{token}_i) = \text{softmax}(W_{\text{class}} h_i + b_{\text{class}})
\]

where $h_i$ is the LayoutLMv3 hidden state for token $i$, and $W_{\text{class}} \in \mathbb{R}^{768 \times 3}$ is a learned projection to the three label classes (Other, Key, Value).

\subsubsection{Training Objective}

Stage 4a uses standard cross-entropy loss over all tokens:

\[
\mathcal{L}_{\text{entity}} = -\sum_{i=1}^{n} \log P(y_i^* \mid \text{token}_i)
\]

where $y_i^* \in \{0, 1, 2\}$ is the ground-truth label for token $i$, and $n$ is the sequence length (padded to 512 tokens).

\subsubsection{Label Generation from Ground Truth}

Ground-truth entity labels are generated from the KVP10k ground truth via bounding box overlap matching:

\begin{enumerate}
    \item For each key-value pair $(K, V)$ in the document's ground truth annotation, retrieve the bounding boxes $\text{bbox}_K$ and $\text{bbox}_V$.
    
    \item For each extracted word $w_i$ with bounding box $\text{bbox}_i$, compute the intersection-over-union (IoU) with both $\text{bbox}_K$ and $\text{bbox}_V$.
    
    \item If $\text{IoU}(\text{bbox}_i, \text{bbox}_K) > 0.5$, label all tokens in word $w_i$ as Key. Similarly for Value.
    
    \item All remaining tokens default to label Other.
\end{enumerate}

This bbox-based alignment avoids text-matching ambiguity and leverages the spatial grounding already present in the KVP10k annotations.

\subsection{Stage 4b: Entity + Relation Linking}
\label{sec:stage4b_entity_relation}

\subsubsection{Architecture}
\label{sec:linker_module}

Stage 4b extends Stage 4a by adding the \texttt{BiaffineLinker} module to score potential key-value relationships. For each (key, value) candidate token pair, the linker computes a scalar score combining a semantic signal and a spatial signal.

\paragraph{Semantic scorer (Stage~4b final).}
The semantic score is a scaled dot-product between projected key and value representations:
\[
    s_{\text{sem}}(i,j) = \frac{\mathbf{k}_i^\top \mathbf{v}_j}{\sqrt{d}}
\]
where $\mathbf{k}_i = W_k h_i$ and $\mathbf{v}_j = W_v h_j$ are linear projections of the LayoutLMv3 hidden states, and $d = 768$ is the hidden dimension. This scoring rule is identical to the attention mechanism in~\cite{vaswani2017attention} and is numerically stable because no learned weight matrix couples the key and value streams in the scoring step.

\textbf{Design note.}
An earlier implementation used \texttt{nn.Bilinear}$(d, d, 1)$, i.e.\ $s = \mathbf{k}^\top W_{\text{bil}} \mathbf{v}$ with a $768 \times 768$ weight matrix. This caused \texttt{link\_loss~=~NaN} at approximately step~1400 of every training run due to exploding gradients: gradients flow back through both $\mathbf{k}$ and $\mathbf{v}$ simultaneously via $W_{\text{bil}}$, destabilising training even under Xavier initialisation. The scaled dot-product formulation eliminates this failure mode at no cost in modelling capacity (commit \texttt{01c3f72}).

\paragraph{Spatial scorer.}
Eight geometric features are computed per (key, value) pair: horizontal and vertical centre displacement $(\Delta x, \Delta y)$, Euclidean distance, angle $\theta = \text{atan2}(\Delta y, \Delta x)$, horizontal and vertical bounding-box alignment overlap ratios, relative area ratio, and relative aspect-ratio ratio. These are encoded by a two-layer MLP with ReLU activations into a 32-dimensional vector.

\paragraph{Final scoring and memory efficiency.}
The semantic scalar and 32-dimensional spatial vector are concatenated and passed through a second MLP to produce the final link score $s_{i,j}$. To avoid materialising the full $n_k \times n_v$ expanded tensor in GPU memory, scoring is performed in chunks of \texttt{LINKER\_CHUNK\_SIZE}~$= 8$ key rows at a time. A hard cap of \texttt{MAX\_LINK\_CANDIDATES}~$= 128$ key and value tokens per page (selected by highest entity-classifier confidence) ensures the worst-case linker matrix is $128 \times 128$, regardless of entity-classifier noise at early training steps.

\subsubsection{Training Objective}

Stage 4b optimizes a weighted combination of entity and relation losses:

\[
\mathcal{L}_{\text{total}} = \mathcal{L}_{\text{entity}} + \lambda \, \mathcal{L}_{\text{relation}}
\]

where $\mathcal{L}_{\text{entity}}$ is the cross-entropy loss from Stage 4a, and $\mathcal{L}_{\text{relation}}$ is binary cross-entropy over predicted relation scores:

\[
\mathcal{L}_{\text{relation}} = -\frac{1}{|\mathcal{B}|} \sum_{b \in \mathcal{B}} \sum_{i,j} \left[ r^*_{i,j} \log \sigma(s_{i,j}) + (1 - r^*_{i,j}) \log(1 - \sigma(s_{i,j})) \right]
\]

Here $r^*_{i,j}$ is the ground-truth relation label (1 if token $i$ is a Key and token $j$ is its corresponding Value, 0 otherwise), $\sigma$ is the sigmoid function, and $|\mathcal{B}|$ is the number of training pages with at least one key-value pair.

\subsubsection{Hyperparameter Sweep}

To explore the interplay between entity and relation learning, we conduct an ablation varying $\lambda \in \{0.5, 1.0, 2.0\}$:

\begin{itemize}
    \item $\lambda = 0.5$: Emphasise entity learning; relation signal is secondary.
    \item $\lambda = 1.0$: Balanced weighting of entity and relation objectives.
    \item $\lambda = 2.0$: Emphasise relation learning; more aggressive link prediction.
\end{itemize}

This sweep determines whether the task is entity-limited (relations are easy given correct entities) or relation-limited (entities are easy but linking is hard).

\subsection{Data Preprocessing and Alignment}
\label{sec:stage4_data_prep}

The Stage 4 pipeline processes raw KVP10k data through several alignment steps:

\subsubsection{Text and Layout Extraction}

KVP10k provides lmdx\_text format: a sequence of extracted text lines, each annotated with pixel-level bounding boxes. Preprocessing extracts:

\begin{enumerate}
    \item \textbf{Word tokens}: Split each lmdx line on whitespace to obtain individual word-level text and bounding boxes.
    \item \textbf{Coordinate normalization}: Scale pixel coordinates from image-native range to $[0, 1000]$ (LayoutLMv3 standard).
    \item \textbf{Image handling}: When real document images are unavailable, create dummy 224\texttimes{}224 white images to satisfy LayoutLMv3Processor requirements.
\end{enumerate}

\subsubsection{LayoutLMv3Processor Configuration}

The LayoutLMv3Processor is configured with \texttt{apply\_ocr=False} to respect pre-extracted text and bounding boxes rather than performing new OCR. The processor:

\begin{enumerate}
    \item Tokenizes word sequences using the RoBERTa tokenizer (subword tokenization).
    \item Aligns subword tokens back to word-level bounding boxes via \texttt{word\_ids()} mapping.
    \item Produces input tensors: \texttt{input\_ids} (token IDs), \texttt{bbox} (normalized coordinates), \texttt{pixel\_values} (image embeddings).
\end{enumerate}

\subsubsection{Token-to-Word Label Alignment}

Ground-truth entity labels are generated at word level but must be aligned to subword tokens for loss computation:

\[
y_{\text{token}} = y_{\text{word}} \quad \text{if } \texttt{word\_ids}[\text{token}] = \text{word}
\]

For special tokens (CLS, SEP, padding) where \texttt{word\_ids} is None, labels are set to 0 (Other) to exclude from loss computation.

\subsubsection{Dataset Format Conversion}
\label{sec:dataset_format_fix}

During Stage 4b development, a mismatch was identified between the raw KVP10k ground-truth annotation format and the binary link-label matrix required by the linker. Specifically, ground-truth key-value relationships are stored as nested reference attributes in the KVP10k JSON; converting these to a dense $[\text{seq\_len} \times \text{seq\_len}]$ binary adjacency matrix required explicit resolution of token-level indices after subword tokenization. Pages with unresolvable references (e.g.\ missing bounding box overlaps) are excluded from link-loss computation but retained for entity-loss computation. Full details are documented in \texttt{STAGE4\_DATASET\_FIX.md}.

\subsection{Training Configuration}
\label{sec:stage4_training_config}

\subsubsection{Optimizer and Learning Rate}
\label{sec:stage4_lr}

Stage 4a uses a single AdamW optimizer with a uniform learning rate across all parameters. Stage 4b uses a \textbf{parameter-group learning rate schedule} to account for the different convergence dynamics of the pre-trained encoder and the randomly-initialised linker:

\begin{itemize}
    \item \textbf{LayoutLMv3 encoder}: $\text{lr}_{\text{enc}} = 2 \times 10^{-5}$ (standard fine-tuning rate for pre-trained transformers)
    \item \textbf{BiaffineLinker (key/value projections, spatial MLP, final scorer)}: $\text{lr}_{\text{link}} = 1 \times 10^{-4}$ (2$\times$ higher, as these layers are trained from random initialisation)
\end{itemize}

Using a single uniform rate for both components caused the linker to under-fit (too slow to learn from random init) while simultaneously destabilising the pre-trained encoder (too fast for fine-tuning). Decoupled learning rates resolved this imbalance.

Both parameter groups share:
\begin{itemize}
    \item Warmup steps: 1000 (5\% of total training steps)
    \item Weight decay: $\lambda_{\text{decay}} = 0.01$
    \item Gradient clipping: $\|\nabla\| \leq 1.0$
    \item Schedule: linear decay after warmup
\end{itemize}

\subsubsection{Numerical Precision}
\label{sec:stage4_precision}

All Stage 4 training is conducted in full \textbf{fp32 precision}. An earlier configuration enabled \texttt{bf16} (bfloat16) mixed-precision training to reduce memory usage. However, \texttt{bf16} caused \texttt{loss~=~NaN} at step~0 of early Stage~4b runs due to underflow in the spatial feature encoder (the geometric feature values, e.g.\ normalised distances and area ratios, fall near the \texttt{bf16} representable range boundary). Reverting to \texttt{fp32} eliminated this issue with negligible impact on memory budget given the chunked linker design.

\subsubsection{Training Stability: NaN Guard}
\label{sec:stage4_nan_guard}

To prevent silent NaN propagation during training (where a NaN loss updates no parameters but corrupts optimiser state), the training loop includes an explicit guard:

\begin{verbatim}
if torch.isnan(loss):
    logger.warning(f"NaN loss at step {step}, skipping batch")
    optimizer.zero_grad()
    continue
\end{verbatim}

This ensures that any residual numerical instability (e.g.\ from degenerate batches with zero key or value candidates) causes an immediate logged skip rather than a silent corruption of the model weights. Together with the Bilinear~$\to$~dot-product fix (Section~\ref{sec:linker_module}) and the fp32 precision decision (Section~\ref{sec:stage4_precision}), this guard makes Stage~4b training robust across all three $\lambda$ configurations.

\subsubsection{Batch Size and Epoch Count}

\begin{itemize}
    \item Batch size: 16 samples per GPU (effective batch size 16 \texttimes{} 4 GPUs = 64 on A100 hardware)
    \item Training epochs: 5 (chosen based on typical BERT fine-tuning convergence on similar-sized tasks)
    \item Validation frequency: Every 100 mini-batches
    \item Early stopping: Patience = 3 epochs (stop if validation F1 does not improve)
\end{itemize}

\subsubsection{Hardware and Time Estimates}

Training is conducted on a single A100 GPU at (estimated):

\begin{itemize}
    \item Stage 4a: \textasciitilde 4 hours for 5 epochs over 5,389 training samples
    \item Stage 4b (single $\lambda$): \textasciitilde 5 hours per run (additional relation scoring overhead)
    \item Total for all 4b experiments ($\lambda \in \{0.5, 1.0, 2.0\}$): \textasciitilde 19 hours
\end{itemize}

\subsection{Inference and Postprocessing}
\label{sec:stage4_inference}

\subsubsection{Stage 4a Inference}

At inference, the model produces token-level probability distributions over label classes. Predictions are:

\begin{enumerate}
    \item Take argmax over predicted label probabilities to assign hard labels.
    \item Aggregate subword token labels back to word level by taking the label of the first subword token of each word.
    \item Extract contiguous spans of Key or Value tokens as final key and value phrases.
\end{enumerate}

\subsubsection{Stage 4b Inference}

Stage 4b additionally computes relation scores and constructs key-value associations:

\begin{enumerate}
    \item Predict entity labels as in Stage 4a.
    \item Compute relation scores $s_{i,j}$ for all (key, value) token pairs.
    \item Extract maximum-weight bipartite matching between Key and Value spans using the relation scores as edge weights.
    \item Return (key phrase, value phrase) pairs along with confidence scores from the linker.
\end{enumerate}

\subsubsection{Bounding Box Prediction}

Both Stage 4a and 4b inherit bounding boxes directly from the preprocessed lmdx\_text, avoiding the need for separate bbox regression. This design choice:

\begin{itemize}
    \item Eliminates hallucinated bounding boxes (common in generative models)
    \item Maintains spatial grounding consistency with ground truth
    \item Simplifies the model (no bbox regression head required)
\end{itemize}

\subsection{Ablation Study Design}
\label{sec:stage4_ablation}

Stage 4 is structured around a series of controlled ablations to isolate the contributions of each component:

\begin{enumerate}
    \item \textbf{Stage 4a (entity-only) vs Stage 4b (entity + relation)}: Measures the performance gain from explicit relation learning. If Stage 4b substantially improves F1, relations are learnable; if Stage 4a is already strong, the task is entity-dominated.
    
    \item \textbf{Relation weight $\lambda$ sweep}: Stage 4b is trained with $\lambda \in \{0.5, 1.0, 2.0\}$ to determine whether the model benefits from emphasising relation predictions ($\lambda > 1$) or entity predictions ($\lambda < 1$).
    
    \item \textbf{Stage 3 vs Stage 4 comparison}: Comparing generative Stage 3 (Mistral-7B, text-only) against discriminative Stage 4 (LayoutLMv3, layout-aware) isolates the contribution of spatial grounding to end-to-end KVP extraction accuracy.
\end{enumerate}

These ablations directly test the hypothesis that layout awareness and spatial grounding improve over text-only generative approaches on the KVP extraction task.

% ------------------------------------------------------------------------------
% ARCHITECTURAL CONSTRAINT NOTE (Section 8.6.1.2)
% Paste into LaTeX_Thesis/chapters/08_discussion.tex under:
%   \subsubsection{Architectural Constraints}
% ------------------------------------------------------------------------------
%
% \subsubsection{Architectural Constraints}
% \label{sec:arch_constraints}
%
% The \texttt{BiaffineLinker} underwent one architectural revision during
% Stage~4b training. The original design used \texttt{nn.Bilinear}$(768, 768, 1)$
% as the semantic scorer, which produced \texttt{link\_loss = NaN} at
% approximately step~1400 across all three $\lambda$ configurations. The root
% cause is that a $768 \times 768$ weight matrix receiving simultaneous gradients
% from both key and value streams is prone to gradient explosion even under Xavier
% initialisation. The fix (commit \texttt{01c3f72}) replaces this with a
% parameter-free scaled dot-product scorer $(\mathbf{k}^\top \mathbf{v}) /
% \sqrt{d}$. All other components and hyperparameters are unchanged.
% Stage~4b results are reported in Section~\ref{sec:stage4b_results}.
%
% Three additional training-stability measures were required:
% (1)~\textbf{fp32 precision}: \texttt{bf16} mixed-precision caused NaN at
%     step~0 due to underflow in spatial feature values; reverted to fp32
%     (see Section~\ref{sec:stage4_precision}).
% (2)~\textbf{Decoupled learning rates}: the linker requires
%     $\text{lr}_{\text{link}} = 10^{-4}$ vs.\ $\text{lr}_{\text{enc}} = 2\times 10^{-5}$
%     for the encoder; a single uniform rate caused linker under-fitting
%     (see Section~\ref{sec:stage4_lr}).
% (3)~\textbf{NaN guard}: explicit \texttt{torch.isnan(loss)} check skips
%     degenerate batches and prevents silent optimiser-state corruption
%     (see Section~\ref{sec:stage4_nan_guard}).
%
% Gradient checkpointing was not enabled: LayoutLMv3 does not reliably support
% the HuggingFace gradient-checkpointing API. Memory was managed via chunked
% linker scoring and the \texttt{MAX\_LINK\_CANDIDATES} cap
% (see Section~\ref{sec:linker_module}).
